Dominance Regions for Rank Two Cluster Algebras

### 2022-05-07 dominance_regions.tex
Minor intro edits
Fixed proof of Lemma 2.3

  When $bc\le3$, the cluster algebra $\cA(b,c)$ will be of finite-type and cluster monomials form its only pointed basis.
  We will therefore assume that $bc\ge4$.
  In this case we give an explicit description of the dominance relation among $\bfg$-vectors.
  Specifically, we show that the $\bfg$-vector $\lambda$ dominates the collection of $\bfg$-vectors of the form $\lambda+(b \alpha_0 ,c \alpha_1)$, $\alpha_0,\alpha_1\in\ZZ$, inside its \emph{dominance region} $\cP_\lambda$.
  %In this case, we compute the possible transitions between any two pointed bases.
  %The collection of $\bfg$-vectors which can appear in the deformation of a pointed basis is the \emph{dominance region}.
  \begin{theorem}
    \label{th:dominance inequalities}
    The dominance region $\cP_\lambda$ pointed at $\lambda=(\lambda_0,\lambda_1)$ is the polygon consisting of those $\mu=(\mu_0,\mu_1)\in\RR^2$ satisfying $\mu_0 \leq \lambda_0, \mu_1 \geq\lambda_1$, and the following inequalities:
    {
      \everymath={\displaystyle}
      \def\arraystretch{2.8}
      \[
        \begin{array}{rcccl}
          0 & \leq & \frac{b c-\sqrt{b c (b c-4)}}{2 b}(\mu_0-\lambda_0)+(\mu_1-\lambda_1) & \leq & \max\left(-c\lambda_0-\frac{b c+\sqrt{b c (b c-4)}}{2b}b\lambda_1,0\right)
          \\
          \min\left(-c\lambda_0,0\right) & \leq & \frac{b c-\sqrt{b c (b c-4)}}{2 b}(\mu_0-\lambda_0)-(\mu_1-\lambda_1) & \leq & 0
          \\
          0 & \leq &  -(\mu_0-\lambda_0)-\frac{b c-\sqrt{b c (b c-4)}}{2 c}(\mu_1-\lambda_1) & \leq & \max\left(\frac{b c+\sqrt{b c (b c-4)}}{2c}c\lambda_0+b\lambda_1,0\right)
          \\
          \min\left(b \lambda_1,0\right) & \leq & (\mu_0-\lambda_0) - \frac{b c-\sqrt{b c (b c-4)}}{2 c} (\mu_1-\lambda_1) & \leq & 0
        \end{array}
      \]
    }
  \end{theorem}

@@

  Given a Laurent polynomial in $\ZZ[x_0^{\pm1},x_1^{\pm1}]$, its \emph{support} is the set of its exponent vectors inside $\ZZ^2$.
  Given a $\bfg$-vector $\lambda$, in the next result we identify a polygon $\cS_\lambda$ whose lattice points are the maximum possible support of a pointed basis element $x_\lambda$. 
  \begin{theorem}
    \label{th:maximum support}
    Let $\lambda=(\lambda_0,\lambda_1)\in\ZZ^2$ be the $\bfg$-vector for a pointed basis element $x_\lambda$.
    \begin{enumerate}
      \item If $\lambda$ lies outside of $\cI$, then the support of $x_\lambda$ is precisely the points of the form $\lambda+(b \alpha_0 ,c \alpha_1)$, $\alpha_0,\alpha_1\in\ZZ$, inside $\cS_\lambda$ given as follows:
        \begin{enumerate}
          \item If $0\le\lambda_0,\lambda_1$, then $\cS_\lambda$ is just the point $\lambda$.
          \item If $\lambda_1 < 0$ and $0\le\lambda_0+b\lambda_1$, then $\cS_\lambda$ is the segment joining $\lambda$ and $(\lambda_0+b\lambda_1,\lambda_1)$.
          \item If $\lambda_0 < 0$ and $0\le\lambda_1$, then $\cS_\lambda$ is the segment joining $\lambda$ and $(\lambda_0,-c\lambda_0+\lambda_1)$.
          \item If $\lambda_0,\lambda_1 < 0$, then $\cS_\lambda$ has corner vertices $\lambda$, $(\lambda_0+b\lambda_1,\lambda_1)$, $\big(\lambda_0+b\lambda_1,-c\lambda_0-(bc-1)\lambda_1\big)$, and $(\lambda_0,-c\lambda_0+\lambda_1)$.
          \item If $\lambda_0 > 0$, $\lambda_0+b\lambda_1 < 0$, and $-c\lambda_0-(bc-1)\lambda_1\le 0$, then $\cS_\lambda$ has corner vertices $\lambda$, $(\lambda_0+b\lambda_1,\lambda_1)$, $\big(\lambda_0+b\lambda_1,-c\lambda_0-(bc-1)\lambda_1\big)$, and $\big((bc+1)\lambda_0-b^2c\lambda_1,-c\lambda_0-(bc-1)\lambda_1\big)$.
          \item If $\lambda_0 > 0$, $\lambda_0+b\lambda_1 < 0$, and $0 < -c\lambda_0-(bc-1)\lambda_1$, then $\cS_\lambda$ has corner vertices $\lambda$, $(\lambda_0+b\lambda_1,\lambda_1)$, $\big(\lambda_0+b\lambda_1,-c\lambda_0-(bc-1)\lambda_1\big)$, and $(0,0)$.
        \end{enumerate}
      \item If $\lambda$ lies inside of $\cI$, then the support of $x_\lambda$ is contained in $\cS_\lambda$ with corner vertices $\lambda$, $(\lambda_0+b\lambda_1,\lambda_1)$, $\big(\lambda_0+b\lambda_1,-c\lambda_0-(bc-1)\lambda_1\big)$, and $\frac{bc+\sqrt{bc(bc-4)}}{2\sqrt{bc(bc-4)}}\big(2\lambda_0+b\lambda_1,-c\lambda_0-(bc-2)\lambda_1\big)$.
        Moreover, there is an element pointed at $\lambda$ whose support is precisely the points of the form $\lambda+(b \alpha_0 ,c \alpha_1)$, $\alpha_0,\alpha_1\in\ZZ$, inside $\cS_\lambda$.
    \end{enumerate}
  \end{theorem}

@@

  \begin{lemma}
    We have
    \[\lim_{i\to\infty} \frac{u_i^-}{u_{i-1}^+}=\frac{bc+\sqrt{bc(bc-4)}}{2b}=\lim_{i\to-\infty} \frac{u_{i-1}^-}{u_i^+};\]
    \[\lim_{i\to\infty} \frac{u_{i-1}^-}{u_i^+}=\frac{bc-\sqrt{bc(bc-4)}}{2b}=\lim_{i\to-\infty} \frac{u_i^-}{u_{i-1}^+}.\]
    Moreover, each of these sequences is monotonic.
  \end{lemma}
  \begin{remark}
    The analogous limits with $+$ and $-$ reversed are obtained from these by interchanging the roles of $b$ and $c$.
  \end{remark}
  \begin{proof}
    For any $i\ne1$, we have
    \[\frac{\big(\sqrt{bc}+\sqrt{bc-4}\big)^i-\big(\sqrt{bc}-\sqrt{bc-4}\big)^i}{\big(\sqrt{bc}+\sqrt{bc-4}\big)^{i-1}-\big(\sqrt{bc}-\sqrt{bc-4}\big)^{i-1}}=\frac{\sqrt{bc}+\sqrt{bc-4}}{1-\left(\frac{\sqrt{bc}-\sqrt{bc-4}}{\sqrt{bc}+\sqrt{bc-4}}\right)^{i-1}}.\]
    It follows that
    \[\lim_{i\to\infty} \frac{u_i^-}{u_{i-1}^+} = \lim_{i\to\infty} \left( \frac{\sqrt{c}}{2\sqrt{b}}\cdot\frac{\sqrt{bc}+\sqrt{bc-4}}{1-\left(\frac{\sqrt{bc}-\sqrt{bc-4}}{\sqrt{bc}+\sqrt{bc-4}}\right)^{i-1}} \right) = \frac{\sqrt{c}}{2\sqrt{b}}\cdot\big(\sqrt{bc}+\sqrt{bc-4}\big),\]
    which is equivalent to the desired expression.
    Similarly, for $i\ne0$ we have
    \[\frac{\big(\sqrt{bc}+\sqrt{bc-4}\big)^{i-1}-\big(\sqrt{bc}-\sqrt{bc-4}\big)^{i-1}}{\big(\sqrt{bc}+\sqrt{bc-4}\big)^i-\big(\sqrt{bc}-\sqrt{bc-4}\big)^i}=\frac{\sqrt{bc}-\sqrt{bc-4}}{4\left(1-\left(\frac{\sqrt{bc}-\sqrt{bc-4}}{\sqrt{bc}+\sqrt{bc-4}}\right)^i\right)},\]
    so that
    \[\lim_{i\to\infty} \frac{u_{i-1}^-}{u_i^+} = \lim_{i\to\infty} \left( \frac{2\sqrt{c}}{\sqrt{b}}\cdot\frac{\sqrt{bc}-\sqrt{bc-4}}{4\left(1-\left(\frac{\sqrt{bc}-\sqrt{bc-4}}{\sqrt{bc}+\sqrt{bc-4}}\right)^{i-1}\right)} \right) = \frac{\sqrt{c}}{2\sqrt{b}}\cdot\big(\sqrt{bc}-\sqrt{bc-4}\big),\]
    which is again equivalent to the desired expression.
    This proves the claim for $i\to\infty$, the cases $i\to-\infty$ follow from these using $u_{-i}^\varepsilon=-u_i^\varepsilon$.

>>>>

  When $bc\le3$, the cluster algebra $\cA(b,c)$ will be of finite-type and cluster monomials form its only pointed basis.
  We will therefore assume that $bc\ge4$.
  In this case we give an explicit description of the dominance relation among $\bfg$-vectors.
  Specifically, we show that the $\bfg$-vector $\lambda$ dominates the collection of $\bfg$-vectors of the form $\lambda+(b \alpha_0 ,c \alpha_1)$, $\alpha_0,\alpha_1\in\ZZ$, inside its \emph{dominance region} $\cP_\lambda$.
  %In this case, we compute the possible transitions between any two pointed bases.
  %The collection of $\bfg$-vectors which can appear in the deformation of a pointed basis is the \emph{dominance region}.
  \begin{theorem}
    \label{th:dominance inequalities}
    The dominance region $\cP_\lambda$ pointed at $\lambda=(\lambda_0,\lambda_1)$ is the polygon consisting of those $\mu=(\mu_0,\mu_1)\in\RR^2$ satisfying $\mu_0 \leq \lambda_0, \mu_1 \geq\lambda_1$, and the following inequalities:
    {
      \everymath={\displaystyle}
      \def\arraystretch{2.8}
      \[
        \begin{array}{rcccl}
          0 & \leq & \frac{b c-\sqrt{b c (b c-4)}}{2 b}(\mu_0-\lambda_0)+(\mu_1-\lambda_1) & \leq & \max\left(-c\lambda_0-\frac{b c+\sqrt{b c (b c-4)}}{2b}b\lambda_1,0\right)
          \\
          \min\left(-c\lambda_0,0\right) & \leq & \frac{b c-\sqrt{b c (b c-4)}}{2 b}(\mu_0-\lambda_0)-(\mu_1-\lambda_1) & \leq & 0
          \\
          0 & \leq &  -(\mu_0-\lambda_0)-\frac{b c-\sqrt{b c (b c-4)}}{2 c}(\mu_1-\lambda_1) & \leq & \max\left(\frac{b c+\sqrt{b c (b c-4)}}{2c}c\lambda_0+b\lambda_1,0\right)
          \\
          \min\left(b \lambda_1,0\right) & \leq & (\mu_0-\lambda_0) - \frac{b c-\sqrt{b c (b c-4)}}{2 c} (\mu_1-\lambda_1) & \leq & 0
        \end{array}
      \]
    }
    %\begin{align}
    %  0 & \leq \frac{b c-\sqrt{b c (b c-4)}}{2 b}(\mu_0-\lambda_0)+(\mu_1-\lambda_1) && \leq \max\left(-c\lambda_0-\frac{b c+\sqrt{b c (b c-4)}}{2b}b\lambda_1,0\right)
    %  \\
    %  \min\left(-c\lambda_0,0\right) & \leq \frac{b c-\sqrt{b c (b c-4)}}{2 b}(\mu_0-\lambda_0)-(\mu_1-\lambda_1) && \leq 0
    %  \\
    %  0 & \leq -(\mu_0-\lambda_0)-\frac{b c-\sqrt{b c (b c-4)}}{2 c}(\mu_1-\lambda_1) && \leq \max\left(\frac{b c+\sqrt{b c (b c-4)}}{2c}c\lambda_0+b\lambda_1,0\right)
    %  \\
    %  \min\left(b \lambda_1,0\right) & \leq (\mu_0-\lambda_0) - \frac{b c-\sqrt{b c (b c-4)}}{2 c} (\mu_1-\lambda_1) && \leq 0
    %\end{align}

  \end{theorem}

@@

  Given a Laurent polynomial in $\ZZ[x_0^{\pm1},x_1^{\pm1}]$, its \emph{support} is the set of its exponent vectors inside $\ZZ^2$.
  Given a $\bfg$-vector $\lambda$, in the next result we identify a polygon $\cS_\lambda$ whose lattice points are the maximum possible support of a pointed basis element $x_\lambda$. 
  \begin{theorem}
    \label{th:maximum support}
    Let $\lambda=(\lambda_0,\lambda_1)\in\ZZ^2$ be the $\bfg$-vector for a pointed basis element $x_\lambda$.
    \begin{enumerate}
      \item If $\lambda$ lies outside of $\cI$, then the support of $x_\lambda$ is precisely the points of the form $\lambda+(b \alpha_0 ,c \alpha_1)$, $\alpha_0,\alpha_1\in\ZZ$, inside $\cS_\lambda$ given as follows:
        \begin{enumerate}
          \item If $0\le\lambda_0,\lambda_1$, then $\cS_\lambda$ is just the point $\lambda$.
          \item If $\lambda_1 < 0$ and $0\le\lambda_0+b\lambda_1$, then $\cS_\lambda$ is the segment joining $\lambda$ and $(\lambda_0+b\lambda_1,\lambda_1)$.
          \item If $\lambda_0 < 0$ and $0\le\lambda_1$, then $\cS_\lambda$ is the segment joining $\lambda$ and $(\lambda_0,-c\lambda_0+\lambda_1)$.
          \item If $\lambda_0,\lambda_1 < 0$, then $\cS_\lambda$ has corner vertices $\lambda$, $(\lambda_0+b\lambda_1,\lambda_1)$, $\big(\lambda_0+b\lambda_1,-c\lambda_0-(bc-1)\lambda_1\big)$, and $(\lambda_0,-c\lambda_0+\lambda_1)$.
          \item If $\lambda_0 > 0$, $\lambda_0+b\lambda_1 < 0$, and $-c\lambda_0-(bc-1)\lambda_1\le 0$, then $\cS_\lambda$ has corner vertices $\lambda$, $(\lambda_0+b\lambda_1,\lambda_1)$, $\big(\lambda_0+b\lambda_1,-c\lambda_0-(bc-1)\lambda_1\big)$, and $\big((bc+1)\lambda_0-b^2c\lambda_1,-c\lambda_0-(bc-1)\lambda_1\big)$.
          \item If $\lambda_0 > 0$, $\lambda_0+b\lambda_1 < 0$, and $0 < -c\lambda_0-(bc-1)\lambda_1$, then $\cS_\lambda$ has corner vertices $\lambda$, $(\lambda_0+b\lambda_1,\lambda_1)$, $\big(\lambda_0+b\lambda_1,-c\lambda_0-(bc-1)\lambda_1\big)$, and $(0,0)$.  Here the point $(0,0)$ and its adjacent open segments are excluded from $\cS_\lambda$.
        \end{enumerate}
      \item If $\lambda$ lies inside of $\cI$, then the support of $x_\lambda$ is contained in $\cS_\lambda$ with corner vertices $\lambda$, $(\lambda_0+b\lambda_1,\lambda_1)$, $\big(\lambda_0+b\lambda_1,-c\lambda_0-(bc-1)\lambda_1\big)$, and $\frac{bc+\sqrt{bc(bc-4)}}{2\sqrt{bc(bc-4)}}\big(2\lambda_0+b\lambda_1,-c\lambda_0-(bc-2)\lambda_1\big)$.
        Moreover, there is an element pointed at $\lambda$ whose support is precisely the points of the form $\lambda+(b \alpha_0 ,c \alpha_1)$, $\alpha_0,\alpha_1\in\ZZ$, inside $\cS_\lambda$.
    \end{enumerate}
  \end{theorem}

@@

  \begin{lemma}
    We have
    \[\lim_{i\to\infty} \frac{u_i^-}{u_{i-1}^+}=\frac{bc+\sqrt{bc(bc-4)}}{2b}=\lim_{i\to-\infty} \frac{u_{i-1}^-}{u_i^+};\]
    \[\lim_{i\to\infty} \frac{u_{i-1}^-}{u_i^+}=\frac{bc-\sqrt{bc(bc-4)}}{2b}=\lim_{i\to-\infty} \frac{u_i^-}{u_{i-1}^+}.\]
    Moreover, each of these sequences is monotonic.
  \end{lemma}
  \begin{remark}
    The analogous limits with $+$ and $-$ reversed are obtained from these by interchanging the roles of $b$ and $c$.
  \end{remark}
  \begin{proof}
    For any $i\ne1$, we have
    \[\frac{\big(\sqrt{bc}+\sqrt{bc-4}\big)^i-\big(\sqrt{bc}-\sqrt{bc-4}\big)^i}{\big(\sqrt{bc}+\sqrt{bc-4}\big)^{i-1}-\big(\sqrt{bc}-\sqrt{bc-4}\big)^{i-1}}=\frac{\big(\sqrt{bc}+\sqrt{bc-4}\big)\cdot\left(1-\left(\frac{\sqrt{bc}-\sqrt{bc-4}}{\sqrt{bc}+\sqrt{bc-4}}\right)^i\right)}{1-\left(\frac{\sqrt{bc}-\sqrt{bc-4}}{\sqrt{bc}+\sqrt{bc-4}}\right)^{i-1}}.\]
    It follows that
    \[\lim_{i\to\infty} \frac{u_i^-}{u_{i-1}^+} = \lim_{i\to\infty} \left( \frac{\sqrt{c}}{2\sqrt{b}}\cdot\frac{\big(\sqrt{bc}+\sqrt{bc-4}\big)\cdot\left(1-\left(\frac{\sqrt{bc}-\sqrt{bc-4}}{\sqrt{bc}+\sqrt{bc-4}}\right)^i\right)}{1-\left(\frac{\sqrt{bc}-\sqrt{bc-4}}{\sqrt{bc}+\sqrt{bc-4}}\right)^{i-1}} \right) = \frac{\sqrt{c}}{2\sqrt{b}}\cdot\big(\sqrt{bc}+\sqrt{bc-4}\big),\]
    which is equivalent to the desired expression.
    Similarly, for $i\ne0$ we have
    \[\frac{\big(\sqrt{bc}+\sqrt{bc-4}\big)^{i-1}-\big(\sqrt{bc}-\sqrt{bc-4}\big)^{i-1}}{\big(\sqrt{bc}+\sqrt{bc-4}\big)^i-\big(\sqrt{bc}-\sqrt{bc-4}\big)^i}=\frac{\big(\sqrt{bc}-\sqrt{bc-4}\big)\cdot\left(1-\left(\frac{\sqrt{bc}-\sqrt{bc-4}}{\sqrt{bc}+\sqrt{bc-4}}\right)^{i-1}\right)}{4\left(1-\left(\frac{\sqrt{bc}-\sqrt{bc-4}}{\sqrt{bc}+\sqrt{bc-4}}\right)^i\right)},\]
    so that
    \[\lim_{i\to\infty} \frac{u_{i-1}^-}{u_i^+} = \lim_{i\to\infty} \left( \frac{2\sqrt{c}}{\sqrt{b}}\cdot\frac{\big(\sqrt{bc}-\sqrt{bc-4}\big)\cdot\left(1-\left(\frac{\sqrt{bc}-\sqrt{bc-4}}{\sqrt{bc}+\sqrt{bc-4}}\right)^{i-1}\right)}{4\left(1-\left(\frac{\sqrt{bc}-\sqrt{bc-4}}{\sqrt{bc}+\sqrt{bc-4}}\right)^i\right)} \right) = \frac{\sqrt{c}}{2\sqrt{b}}\cdot\big(\sqrt{bc}-\sqrt{bc-4}\big),\]
    which is again equivalent to the desired expression.
    This proves the claim for $i\to\infty$, the cases $i\to-\infty$ follow from these using $u_{-i}^\varepsilon=-u_i^\varepsilon$.
